Data Preparation

### 2021-07-21 x0_data-preparation.Rmd
geocoding for data cleaning and ADEP-ADES assignment

#' utility function to coerce

@@

prep_base_data <- function(.fn = file_names, ...){
ds <- read_osn_covid(.fn)
ds <- ds %>% append_ad_ctry()

ds_ok  <- ds %>% filter(!is.na(ADEP), !is.na(ADES))
ds_nok <- ds %>% filter( is.na(ADEP) | is.na(ADES))

ds <- ds %>% append_region()

@@

# write out base data
my_fn <- gsub(pattern = "-raw", replacement = "", x = .fn)
my_fn <- gsub(pattern = "flightlist_", replacement = "base_osn_", x = my_fn)
my_fn <- gsub(pattern = "\\.gz", replacement = "", x = my_fn)
write_csv(ds, file = my_fn)
message(paste0("run completed for ", my_fn))
}
```

>>>>

#' utility function to coerce
pts_latlon_sf <- function(.df ,.lat, .lon, .drop=FALSE){
  check <- nrow(.df)
  df <- .df %>%
    filter(!is.na(.data[[.lat]]))%>%    # index to character vector ?!!
    sf::st_as_sf(coords = c(.lon, .lat), crs=4326, remove = .drop)

  if(check > nrow(df)) warning(paste0("\nremoved ", check - nrow(df), " flights with NA in lat/lon."))

#' identify territory based on lat/lon
geo_sfpts <- function(.sfpts, .shape){
  pts <- .sfpts %>%
    sf::st_join(.shape)
  return(pts)
}

@@

prep_base_data <- function(.fn = file_names, ...){
ds <- read_osn_covid(.fn)
ds <- ds %>% append_ad_ctry() 
ds <- ds %>% append_region()

ds_ok  <- ds %>% filter(!is.na(ADEP), !is.na(ADES)) %>% select(ADEP:CALL)
ds_nok <- ds %>% filter( is.na(ADEP) | is.na(ADES))

ds_nok_adep <- ds_nok %>% filter(is.na(ADEP)) %>% 
    pts_latlon_sf("LAT1","LON1") %>% 
    geo_sfpts(eez_land) %>% 
    mutate(ISO2 = countrycode(ISO_TER1, origin = "iso3c", destination = "iso2c")
           ,ADEP     = case_when(ALT1 <= 5000 ~ 'XLOX', ALT1 > 5000 ~ 'XHIX', TRUE ~ as.character(NA))
           ,ADEP_CTRY= case_when( ADEP == "XHIX" ~ "Other"
                                 ,ADEP == "XLOX" & !is.na(ISO2) ~ ISO2
                                 ,ADEP == "XLOX" & is.na(ISO2) ~ ADES_CTRY
                                 ,TRUE ~ as.character(NA))
           )

ok <- ds_nok_adep %>% filter(!is.na(ADES))  # check there are a few apts with no country
nok<- ds_nok_adep %>% filter( is.na(ADES)) %>%
  sf::st_drop_geometry() %>% rename(ISO3_ADEP = ISO_TER1, ISO2_ADEP = ISO2) %>%
  pts_latlon_sf("LAT2","LON2") %>% 
  geo_sfpts(eez_land) %>% 
    mutate(ISO2 = countrycode(ISO_TER1, origin = "iso3c", destination = "iso2c")
           ,ADES     = case_when(ALT2 <= 5000 ~ 'XLOX', ALT2 > 5000 ~ 'XHIX', TRUE ~ as.character(NA))
           ,ADES_CTRY= case_when( ADES == "XHIX" ~ "Other"
                                 ,ADES == "XLOX" & !is.na(ISO2) ~ ISO2
                                 ,ADES == "XLOX" & is.na(ISO2) ~ ADEP_CTRY
                                 ,TRUE ~ as.character(NA))
    )

ok <- ok  %>% sf::st_drop_geometry() %>% select(ADEP:CALL)
nok<- nok %>% sf::st_drop_geometry() %>% select(ADEP:CALL)

ds_nok_adep <- bind_rows(ok, nok) %>% append_region()

ds_nok_ades <- ds_nok %>% filter(!is.na(ADEP), is.na(ADES)) %>%
  pts_latlon_sf("LAT2","LON2") %>% 
  geo_sfpts(eez_land) %>% 
    mutate(ISO2 = countrycode(ISO_TER1, origin = "iso3c", destination = "iso2c")
           ,ADES     = case_when(ALT2 <= 5000 ~ 'XLOX', ALT2 > 5000 ~ 'XHIX', TRUE ~ as.character(NA))
           ,ADES_CTRY= case_when( ADES == "XHIX" ~ "Other"
                                 ,ADES == "XLOX" & !is.na(ISO2) ~ ISO2
                                 ,ADES == "XLOX" & is.na(ISO2) ~ ADEP_CTRY
                                 ,TRUE ~ as.character(NA))
    ) %>% 
  sf::st_drop_geometry() %>% select(ADEP:CALL) %>% 
  append_region()

ds_nok <- bind_rows(ds_nok_adep, ds_nok_ades)   # verify number of flights

#--------- combine again
ds <- bind_rows(ds_ok, ds_nok)

#-------------------

@@

# write out base data
my_fn <- gsub(pattern = "-raw", replacement = "", x = .fn)
my_fn <- gsub(pattern = "flightlist_", replacement = "base_osn_2_", x = my_fn)
my_fn <- gsub(pattern = "\\.gz", replacement = "", x = my_fn)
write_csv(ds, file = my_fn)
message(paste0("run completed for ", my_fn))
}
```
